---
title: "Data Wrangling"
author: "Bill Lang"
date: "3/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


Pullin the data in. This was used by manual downloading the data to my github but could have been down with more elegant use of the WRPDC API and the SQL functionality of Rstuido (or just python). Regardless, the data is read in and merged into a single dataset to be used for the regression.

```{r}
houseOwnership <- read.csv(url("https://raw.githubusercontent.com/bill-spec/Health-and-Housing/master/data/housingrentown%20(1).csv?token=ANWDS6XNV7UDBU2K74RWHLS6NBZE2"))

houseOwnership <- houseOwnership %>% 
    mutate(tract = as.integer(tract*100))

housingPoor <- read.csv(url("https://raw.githubusercontent.com/bill-spec/Health-and-Housing/master/data/poorhousingconditions.csv?token=ANWDS6VKUI7U6SSV7CDNC4K6NBZCS"))
housingVacant <- read.csv(url("https://raw.githubusercontent.com/bill-spec/Health-and-Housing/master/data/vacantpropusps2016q2.csv?token=ANWDS6UB3JRTY4OQFQPUJN26NBZG6"))

houseOwnership
housingPoor  #tract
housingVacant

housing <- addByString(houseOwnership, housingVacant , "tract")

housing <- housing[,-6] #Removing Geocoding column.

housing

```


```{r}

childrenLead <- read.csv(url("https://raw.githubusercontent.com/bill-spec/Health-and-Housing/master/data/Children%20Lead.csv?token=ANWDS6RTLZTONJHOIEG4R7S6NBZJM"))
childrenLead <- tractToInt(childrenLead)

anxiety <- read.csv(url("https://raw.githubusercontent.com/bill-spec/Health-and-Housing/master/data/Anxiety.csv?token=ANWDS6TUWZKJZPYUNHLA2S26NBZL2"))
depression <- read.csv(url("https://raw.githubusercontent.com/bill-spec/Health-and-Housing/master/data/Depression.csv?token=ANWDS6WPEYI3WKOME57EGNK6NBZRK"))
diabetes <- read.csv(url("https://raw.githubusercontent.com/bill-spec/Health-and-Housing/master/data/diabetes_all_2016.csv?token=ANWDS6WUFHOZ7UV5XUCKACC6NBZPQ"))
hypertension <- read.csv(url("https://raw.githubusercontent.com/bill-spec/Health-and-Housing/master/data/Hypertension.csv?token=ANWDS6XTPSMERYEVNKCGW2S6NBZQS"))


childrenLead
anxiety
depression
diabetes
hypertension
#asthmaUse

predictors <- addByString(childrenLead, anxiety, "tract")
predictors <- addByString(predictors, depression, "tract")
predictors <- addByString(predictors, diabetes, "tract")
predictors <- addByString(predictors, hypertension, "tract")
#predictors <- addByString(predictors, asthmaUse, "tract")
predictors <- predictors %>% dplyr::select(-GEOID10)

predictors
```


Removal of any probalmatic whitespace. 

```{r}
reg <- addByString(housing, predictors, "tract")

reg$TotalPopulation <- as.numeric(gsub(",","",reg$TotalPopulation))
reg$RenterOccupied <- as.numeric(gsub(",","",reg$RenterOccupied))
reg$OwnedMortgage <- as.numeric(gsub(",","",reg$OwnedMortgage))
reg$OwnedFree <- as.numeric(gsub(",","",reg$OwnedFree))

reg1 <- reg %>% 
  mutate(
              totalPopulation = as.integer(as.character(TotalPopulation)),
              ownedMortgage = as.integer(as.character(OwnedMortgage)),
              ownedFree = as.integer(as.character(OwnedFree)),
              renterOccupied = as.integer(as.character(RenterOccupied)),
              vacantProportion = as.double(as.character(percent.of.vacant.properties)),
              residentialVacancies = as.integer(as.character(ResidentialVacancies)),
              residentialBuildings = as.integer(as.character(ResidentalBuildings))
              ) %>% select(-TotalPopulation,-OwnedMortgage,-OwnedFree,-RenterOccupied,-percent.of.vacant.properties)

reg1

```



Cleaning Lead Levels 

40500 and 40600 (Frazer and Fawn twps) are missing data for the 5 year avearge. Averaging their adjacent regions we can find replacemnts. 

459202 is pennsbury Village and we will say is 1.  

or we call tell all three of these to be zero.
or we can remove themm all together. 



```{r}
#For now we can say they are 0

reg1[6,13] <- 0
reg1[7,13] <- 0
reg1[202,13] <- 0

reg1
```


Mutating Columns to make them per-population

```{r}

#Anxiety, Depression, and tenure of housing 
#Anxiety and Depression it is the proportion of those who are currently 
#taking medication for it. 

reg1 <- reg1 %>% 
  mutate(anxietyProportion = XPAN/XPAD) %>% 
  mutate(anxietyMaleProportion = XMAN/XMAD) %>% 
  mutate(anxietyFemaleProportion = XWAN/XWAD) %>% 
  mutate(depressionProportion = EPAN/EPAD) %>% 
  mutate(depressionMaleProportion = EMAN/EMAD) %>% 
  mutate(depressionFemaleProportion = EWAN/EWAD) %>% 
  mutate(freeProportion = ownedFree/totalPopulation) %>% 
  mutate(mortgageProportion = ownedMortgage/totalPopulation) %>% 
  mutate(rentProportion = renterOccupied/totalPopulation) 

#Hypertension
#Hyper 2 is both diagnosed and medicated.

reg1 <- reg1 %>% 
  mutate(hyperProportion = TPAN/TPAD) %>% 
  mutate(hyperTreatedProportion = TPAN2/TPAD) %>% 
  mutate(hyperFemaleProportion = TWAN/TWAD) %>% 
  mutate(hyperMaleProportion = TMAN/TMAD) %>% 
  mutate(hyperTreatedMaleProportion = TMAN2/TMAD) %>% 
  mutate(hyperTreatedFemaleProportion = TWAN2/TWAD)


#diabetes 

reg1 <- reg1 %>% 
  mutate(diabetesProportion = BPAN/BPAD) %>% 
  mutate(diabetesTreatedProportion = BPAN2/BPAD) %>% 
  mutate(diabetesFemaleProportion = BWAN/BWAD) %>% 
  mutate(diabetesMaleProportion= BMAN/BMAD) %>% 
  mutate(diabetesTreatedMaleProportion = BMAN2/BMAD) %>% 
  mutate(diabetesTreatedFemaleProportion = BWAN2/BWAD)

head(reg1)
```

Counting the population we can see that this encompases most of Allegehny County

```{r}
reg1 %>% arrange(desc(totalPopulation))

reg1 %>% 
  summarise(
  count = sum(totalPopulation)
)
```

Large Dataset

```{r}
regularData <- reg1[,-(9:38)]
regularData <- na.omit(regularData)

regularData <- regularData %>% remove_rownames %>% column_to_rownames(var="tract")
dataSetFull <- cbind(regularData, "Observation.Number" = c(1:380))
dataSetFull
cor(dataSetFull)
```

Small Dataset

```{r}
dataSet <- dataSetFull[,-c(1:6,9:11,13:14,16:17,19:22,24,26:30,32:35)]
dataSet <- dataSet[,c(6,4,5,7,8,1,2,3,9)]
dataSet
```

```{r}
dataSet
```


```{r}
write.csv(dataSetFull, file = "C:/Users/Billy/Desktop/1223 Dataset/Full_Data_Set.csv")
write.csv(dataSet, file = "C:/Users/Billy/Desktop/1223 Dataset/Used_Data_Set.csv")
```
